### 2017-04-05 draft1.tex

% !TeX root = RJwrapper.tex
\title{true}
\author{by Ross Jacobucci}

\maketitle

\abstract{%
The \textbf{regsem} package in \textbf{R}, an implementation of
regularized structural equation modeling (RegSEM; Jacobucci, Grimm, and
McArdle 2016), was recently developed with the goal of incorporating
various forms of penalized likelihood estimation in a broad array of
structural equations models. The forms of regularization include both
the \textit{ridge} (Hoerl and Kennard 1970) and the least absolute
shrinkage and selection operator (\textit{lasso}; Tibshirani 1996),
along with sparser extensions. RegSEM is particularly useful for
structural equation models that have a small parameter to sample size
ratio, as the addition of penalties can reduce the complexity, thus
reducing the bias of the parameter estimates. The paper covers the
algorithmic details and an overview of the use of \textbf{regsem} with
the application of both factor analysis and latent growth curve models.
}

\section{Introduction}\label{introduction}

The desire for simplicity in model structure comes by many names,
including simple structure (Thurstone 1935), variable complexity (Browne
2001), parsimony (Raykov and Marcoulides 1999; Marsh and Hau 1996),
``sparse loadings'' in the context of principal components analysis
(Zou, Hastie, and Tibshirani 2006), and lastly, ``sparsistency'',
denoting that all parameters in a sparse model that are zero are
correctly estimated as zero with probability tending to one (Lam and Fan
2009). The goal is to accurately and efficiently estimate a model that
is parsimonious in allowing users to easily interpret the model's
representation of reality. In the context of latent variables, reducing
the complexity of models can come in many forms: selecting among
multiple predictors of a latent variable, simplifying factor structure
by removing cross-loadings, determining whether the addition of
nonlinear terms are necessary in longitudinal models, and many others.

\begin{figure}
    \centering
    \includegraphics[width=.5\linewidth]{figs/growth_fig}
    \caption{Growth curve model with 10 predictors of both the intercept and slope}
\end{figure}

As a simple running example, Figure 1 depicts a linear latent growth
curve model (e.g. Meredith and Tisak 1990) with four time points and ten
predictors for a simulated dataset. In this, a researcher may want to
test this model, but may only have a relatively small sample size
(e.g.~80). There are 29 estimated parameters in this model, resulting in
a estimated parameter to sample size ratio far below even the most
liberal recommendations (e.g.~10:1 parameters to sample size; Kline
2015). In lieu of finding additional respondents, reducing the number of
parameters estimated is one effective strategy for reducing bias.
Specifically, the 20 estimated regressions from \textit{c1-c10} could be
reduced to a number that makes the ratio of the parameters estimated to
sample size more reasonable. To explore this further, the next section
provides an overview of regularization, and how different forms can be
used to perform variable selection across a broad range of models.

\section{Regularization}\label{regularization}

Although a host of methods exist to perform variable selection, the use
of regularization has seen a wide array of application in the context of
regression, and more recently, in areas such as graphical modeling, as
well as a host of others. The two most common procedures for
regularization in regression are the \textit{ridge} (Hoerl and Kennard
1970) and the least absolute shrinkage and selection operator
(\textit{lasso}; Tibshirani 1996); however, there are various
alternative forms that can be seen as subsets or generalizations of
these two procedures. Given an outcome vector \textit{y} and predictor
matrix \(X \in {R}^{n \times p}\) , ridge estimates are defined as

\[\tag{1}
\hat{\beta}^{ridge}= argmin \Big\{ \sum_{i=1}^{N} (y_{i} = \beta_{0} - \sum_{j=1}^{p}x_{ij} \beta_{j})^{2}  + \lambda \sum_{j=1}^{p} \beta_{j}^{2}\Big\},
\]

where \(\beta_{0}\) is the intercept, \(\beta_{j}\) is the coefficient
for \(x_{j}\), and \(\lambda\) is the penalty that controls the amount
of shrinkage. Note that when \(\lambda = 0\), Equation 3 reduces to
ordinary least squares regression. As \(\lambda\) is increased, the
\(\beta\) parameters are shrunken towards zero. The lasso estimates are
defined as

\[\tag{2}
\hat{\beta}^{lasso}= argmin \Big\{ \sum_{i=1}^{N} (y_{i} = \beta_{0} - \sum_{j=1}^{p}x_{ij} \beta_{j})^{2}  + \lambda \sum_{j=1}^{p}|\beta_{j}|\Big\}.
\]

In lasso regression, the \(l_{1}\)-norm is used, instead of
\(l_{2}\)-norm as in ridge, which also shrinks the \(\beta\) parameters,
but additionally drives the parameters all the way to zero, thus
performing a form of subset selection.

In the context of our example depicted in Figure 1, to use lasso
regression to select among the covariates, the growth model would need
to be reduced to two factor scores, which neglects both the relationship
between both the slope and intercept, reducing both to independent
variables. Particularly in models with a greater number of latent
variables, this becomes increasingly problematic. A method that keeps
the model structure, while allowing for penalized estimation of specific
parameters is regularized structural equation modeling (RegSEM;
Jacobucci, Grimm, and McArdle 2016). RegSEM adds a penalty function to
the traditional maximum likelihood estimation (MLE) for structural
equation models (SEMs). The maximum likelihood cost function for SEMs
can be written as

\[\tag{3}
F_{ML}=log(\left|\Sigma\right|)+tr(C*\Sigma^{-1})-log(\left|C\right|)- p.
\]

where \(\Sigma\) is the model implied covariance matrix, \(C\) is the
observed covariance matrix, and \(p\) is the number of estimated
parameters. RegSEM builds in an additional element to penalize certain
model parameters yielding

\[\tag{4}
F_{regsem} = F_{ML} + \lambda P(\cdot)
\]

where \(\lambda\) is the regularization parameter and takes on a value
between zero and infinity. When \(\lambda\) is zero, MLE is performed,
and when \(\lambda\) is infinity, all penalized parameters are shrunk to
zero. \(P(\cdot)\) is a general function for summing the values of one
or more of the model's parameter matrices. Two common forms of
\(P(\cdot)\) include both the lasso (\(\| \cdot \|_{1}\)), which
penalizes the sum of the absolute values of the parameters, and ridge
(\(\| \cdot \|_{2}\)), which penalizes the sum of the squared values of
the parameters.

In our example, the twenty regression parameters from the covariates to
both the intercept and slope would be penalized. Using lasso penalties,
the absolute value of these twenty parameters would be summed and after
being multiplied by the penalty \(\lambda\), added to equation 4,
resulting in:

\[\tag{5}
F_{lasso} = F_{ML} + \lambda * \left\|  \begin{matrix}  
c1 \xrightarrow[]{} i\\
c2\xrightarrow[]{}i\\
\vdots \\
c10\xrightarrow[]{}i\\
c1\xrightarrow[]{}s\\
\vdots \\
c10\xrightarrow[]{}s\\
\end{matrix}  \right\|_{1}
\]

Although the fit of the model is easily calculated given a set of
parameter estimates, traditional optimization procedures for SEM cannot
be used given the non-differentiable nature of lasso penalties, and as
detailed later, sparse extensions.

\subsection{Optimization}\label{optimization}

One method that has become popular for optimizing penalized likelihood
method is that of proximal gradient descent (e.g.~p.~104 in Hastie,
Tibshirani, and Wainwright 2015). In comparison to one-step procedures
common in SEM optimization, that only involve a method for calculating
the step size and the direction (typically using the gradient and an
approximation of the Hessian), proximal gradient descent can be
formulated as a two-step procedure. With a stepsize of \(s^{t}\) and
parameters \(\theta^{t}\) at iteration \textit{t}:

\begin{enumerate}
    \item First, take a gradient step size $z = \theta^{t} - s^{t} \nabla g(\theta^{t})$.
    \item Second, perform elementwise soft-thresholding $\theta^{t+1} = S_{s^{t} \lambda}(z)$.
\end{enumerate}

where \(S_{s^{t} \lambda}(z)\) is the soft-thresholding operator (Donoho
1995) used to overcome non-differentiability of the lasso penalty at the
origin:

\begin{equation}
S_{s^{t} \lambda}(z_{j}) = sign(\theta_{j})(|\theta_{j}|-s^{t} \lambda)_{+}
\end{equation}

where \((x)_{+}\) is shorthand for max(x,0) and \(s^{t}\) is the step
size. Henceforth, \(\lambda\) is assumed to encompass both the penalty
and the step size \(s^{t}\). This procedure is only used to update
parameters that are subject to penalty. Non-penalized parameters are
updated only using step 1 from above.

In contrast to only updating one parameter at a time in a
coordinate-wise fashion, for RegSEM the optimization steps can be
divided by both the \textit{A} and \textit{S} matrices (see the
Implementation section for more description on the RAM matrices). This
block-wise gradient descent manifests itself as:

\begin{minipage}{\textwidth}
    %   \renewcommand\footnoterule{}    
    \begin{algorithm}[H]
        \begin{algorithmic}[1]
            %   \Procedure{RegSEM Block Coordinate Descent for Lasso}{}
            %   \footnotetext{Note: $A(pen)$ refers only to the penalized parameters, $vec()$ concatenates both vectors}
            \State Generate starting values for $\theta_{t}$
            \State Calculate initial fit $F_{t}$
            \State Set step size $s$. 0.1 works well at this time.
            \State Set tolerance (tol). e.g. 1e-6
            \While {$|F_{t} - F_{t+1}| > tol$}
            \State Calculate gradient for A:$ \nabla (A) =:  \frac{\partial A}{\partial \theta_{t}}$
            \State $ \theta_{t+1^{*},A} =: \theta_{t,A} - s \nabla(A)$
            \State  Update penalized parameters: $\theta_{t+1^{*},A(pen)} =: S_{s \lambda}(\theta_{t+1^{*},A(pen)})$
            \State Calculate gradient for S:$ \nabla (S) =:  \frac{\partial S}{\partial \theta_{t+1^{*}}}$ 
            \State Update S parameters: $ \theta_{t+1^{*},S} =: \theta_{t,S} - s \nabla(S)$
            \State $\theta_{t+1} =: \theta_{t+1^{*}}$ %vec(\theta_{t+1^{*},A},\theta_{t+1^{*},S})$
            \State Update $S_{t+1}, A_{t+1}$
            \State $ \Sigma_{t+1} = F(I-A_{t+1})^{-1}S_{t+1}(I-A_{t+1})^{-T}F^{T}$
            \State $ F_{t+1} = F_{ML}(\Sigma_{t+1},C) + \lambda \| A_{t+1}(pen) \|$
            \EndWhile
            %   \EndProcedure
        \end{algorithmic}
        \caption{RegSEM Block Coordinate Descent}
        \label{alg:seq}
    \end{algorithm}
\end{minipage}

where step 13 is the calculation of the implied covariance matrix using
RAM matrices and step 14 calculates the fit of the model with penalizing
parameters in the \textit{A} matrix. Note that parameters from the
\textit{S} matrix can also be penalized, however, this is much less
common. This algorithm can be described as first order proximal block
coordinate descent. For some SEM models, using block updates has been
found to work better than standard gradient descent with the same soft
thresholding of penalized parameters.

\section{Types of Penalties}\label{types-of-penalties}

Outside of both ridge and lasso penalties, a host of additional forms of
regularization exist.

\subsection{Elastic Net}\label{elastic-net}

Most notably, the elastic net (Zou and Hastie 2005) encompasses both the
ridge and lasso, reaching a compromise between both through the addition
of an additional parameter \(\alpha\), manifesting itself as

\[
P_{enet}(\theta_{j}) = (1-\alpha)\| \theta_{j} \|_{2} + \alpha\| \theta_{j} \|_{1}
\]

with a soft-thresholding update of \[
S(\theta_{j})= 
\begin{cases}
0,&  |\theta_{j}| < \alpha\lambda\\
\frac{sgn(\theta_{j})(|\theta_{j}|-\alpha\lambda)}{1+(1-\alpha)\lambda},              & |\theta_{j}|\geq\alpha\lambda
\end{cases}
\]

When \(\alpha\) is zero, ridge is performed, and conversely when
\(\alpha\) is 1, lasso regularization is performed. This method
harnesses the benefits of both methods, particularly when variable
selection is warranted (lasso), but there may be collinearity between
the variables (ridge).

\subsection{Adaptive Lasso}\label{adaptive-lasso}

In using lasso penalties, difficulties emerge when the scale of
variables differ dramatically. By only using one value of \(\lambda\),
this can add appreciable bias to the resulting estimates (e.g. Fan and
Li 2001). One method proposed for overcoming this limitation is the
adaptive lasso (Zou 2006). Instead of penalizing parameters directly,
each parameter is scaled by the un-penalized estimated (MLE parameter
estimates in SEM). The adaptive lasso results in: \[
F_{alasso} = F_{ML} + \lambda \| \theta_{ML}^{-1} * \theta_{pen} \|_{1}
\]

with, following the same form for the lasso, the soft-thresholding
update is:

\[
S(\theta_{j})= sign(\theta_{j})(|\theta_{j}|-\frac{\lambda}{2|\theta_{j}|})_{+}
\]

In this, larger penalties are given for non-significant (smaller)
parameters, limiting the bias in estimating larger, significant
parameters. Note that one limitation of this approach for SEM models is
that the model needs to be estimable with MLE. Particularly for models
with large numbers of variables, in relation to sample size, this may
not be possible.

\subsection{Sparse Extensionse}\label{sparse-extensionse}

\begin{figure}
    \centering
    \includegraphics[width=.5\linewidth]{figs/penalties}
    \caption{Comparison of types of penalties with $\lambda=0.5$}
\end{figure}

Two additional penalties that overcome some of the deficiencies of the
lasso, producing sparser solutions, include the smoothly clipped
absolute deviation penalty (SCAD; Fan and Li 2001) and the minimax
concave penalty (MCP; Zhang 2010). In comparison to the lasso, both the
SCAD and MCP have much smaller penalties for large parameters, where the
amount of penalty for small penalties is similar to the lasso, as is
evident in Figure 2.

The SCAD takes the form of:

\[
pen_{\lambda,\gamma}(\theta_{j}) = \lambda \big\{I(\theta_{j}\leq\lambda) + \frac{(\gamma \lambda-0)_{+}}{(\gamma-1)\lambda}I(\theta_{j}>\lambda)\big\}
\]

with a soft-thresholding update of \[
S(\theta_{j})= 
\begin{cases}
S(\theta_{j},\lambda),&  |\theta_{j}| \geq 2\lambda\\
\frac{\gamma-1}{\gamma-2}S(\theta_{j},\frac{\lambda\gamma}{\gamma-1}),              & 2\lambda < |\theta_{j}|\leq\alpha\lambda\\
\theta_{j} & |\theta_{j}| > \lambda \gamma
\end{cases}
\]

for \(\gamma > 2\). As the the penalty in equation 11 is non-convex (as
is the MCP), this makes the computation more difficult. However, in the
context of SEM this can be seen as less problematic, as equation 3 is
also non-convex.

Additionally, the MCP takes the form of:

\[
pen_{\lambda,\gamma}(\theta_{j}) = \lambda\bigg(|\theta_{j}|-\frac{\theta_{j}^{2}}{2\lambda\gamma}\bigg)I(|\theta_{j}|<\lambda\gamma) +\frac{\lambda^{2}\gamma}{2}I(|\theta_{j}|\geq \lambda\gamma)
\]

with a soft-thresholding update of \[
S(\theta_{j})= 
\begin{cases}
\frac{\gamma}{\gamma-1}S(\theta_{j},\lambda),&  |\theta_{j}| \leq \lambda\gamma\\
\theta_{j} & |\theta_{j}| > \lambda \gamma
\end{cases}
\]

for \(\gamma > 0\). As seen in Figure 2, this results in similar amount
of shrinkage for smaller estimates in comparison to the SCAD, however,
less for larger estimates. For both the SCAD and MCP, both the
\(\gamma\) and \(\lambda\) parameters are used as hyper-parameters. This
involves testing models over a two-dimensional array of parameters,
however, in \textbf{regsem}, \(\gamma\) is by default fixed to 3.7 per
Fan and Li (2001).

\section{Implementation}\label{implementation}

RegSEM is implemented as the \textbf{regsem} package (Jacobucci 2017) in
the \textbf{R} statistical environment (R Core Team 2017). To estimate
the maximum likelihood fit of the model, \textbf{regsem} uses
\textit{Reticular Action Model} (RAM; McArdle and McDonald 1984; McArdle
2005) notation to derive an implied covariance matrix. The parameters of
each SEM are translated into three matrices: the \textit{filter}
(\textit{F}), the \textit{asymmetric} (\textit{A}; directed paths;
e.g.~factor loadings or regressions), and the \textit{symmetric}
(\textit{S}; undirected paths; e.g.~covariances or variances). See
Jacobucci, Grimm, and McArdle (2016) for more detail on RAM notation and
its application to RegsEm.

Syntax for using the \textbf{regsem} is based on the \textbf{lavaan}
package (Rosseel 2012) for structural equation models. \textbf{lavaan}
is a general SEM software program that can fit a wide array of models
with various estimation methods. To use \textbf{regsem}, the user has to
first fit the model in \textbf{lavaan}. Note that particularly in cases
that the number of variables is larger than the sample size, the model
in lavaan does not need to converge, let alone run. In this case, the
\textbf{do.fit=FALSE} argument in lavaan can be used. As a canonical
example, below is the code for a confirmatory factor analysis model with
one latent factor and nince indicators from the Holzinger and Swineford
(1939) dataset.

\begin{Schunk}
\begin{Sinput}
library(lavaan)
mod <- "
f1 = ~ NA*x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9
f1~~1*f1
"
out <- cfa(mod,HolzingerSwineford1939)
\end{Sinput}
\end{Schunk}

After a model is run in lavaan, using \textbf{lavaan()} or any of the
wrapper functions for fitting a model (i.e. \textbf{sem()},
\textbf{cfa()}, or \textbf{growth()}), the object is then used by the
regsem package to translate the model into RAM notation and run using
one of three functions: \textbf{regsem()}, \textbf{multi_optim()}, or
\textbf{cv_regsem()}. The \textbf{regsem()} function runs a model with
one penalty value, whereas \textbf{multi_optim()} does the same but
allows for the use of random starting values. However, the main function
is \textbf{cv_regsem()} as this not only runs the model, but runs it
across a vector of varying penalty values. For instance in the above
one-factor model, each of the factor loadings can be tested with lasso
penalties to determine whether each indicator is a necessary component
of the latent factor:

\begin{Schunk}
\begin{Sinput}
library(regsem)
extractMatrices(out)["A"]
out.reg <- cv_regsem(out, type="lasso", 
                    pars_pen=c(1:9),n.lambda=15,jump=.05)
\end{Sinput}
\end{Schunk}

In this, the function \textbf{extractMatrices()} allows the user to
examine at how the lavaan model is translated into RAM matrices.
Further, by looking at the \textit{A} matrix, the parameter numbers
corresponding to the factor loadings of interest for regularization can
be identified. For this model, the factor loadings represent parameter
numbers one through nine, of which we pass directly to the
\textbf{pars_pen} argument of the \textbf{cv_regsem()} function (if
\textbf{pars_pen=NULL} then all directed effects, outside of intercepts,
are penalized). Additionally, we pass the arguments of how many values
of penalty we want to test (\textbf{n.lambda=15}), how much the penalty
should increase for each model (\textbf{jump=.05}), and finally that
lasso estimation is used (\textbf{type="lasso"}).

The \textbf{out.reg} object contains two components,
\textbf{out.reg[[1]]} has the parameter estimates for each of the 15
models,

\begin{Schunk}
\begin{Sinput}
head(round(out.reg[[1]],2),5)
\end{Sinput}
\begin{Soutput}
#>      f1 -> x1 f1 -> x2 f1 -> x3 f1 -> x4 f1 -> x5 f1 -> x6 f1 -> x7
#> [1,]     0.51     0.26     0.25     0.98     1.08     0.92     0.20
#> [2,]     0.42     0.17     0.16     0.90     0.99     0.83     0.11
#> [3,]     0.00     0.00     0.00     0.00     0.00     0.00     0.00
#> [4,]     0.00     0.00     0.00     0.00     0.00     0.00     0.00
#> [5,]     0.00     0.00     0.00     0.00     0.00     0.00     0.00
#>      f1 -> x8 f1 -> x9 x1 ~~ x1 x2 ~~ x2 x3 ~~ x3 x4 ~~ x4 x5 ~~ x5
#> [1,]     0.20     0.31      1.1     1.31     1.21     0.38     0.49
#> [2,]     0.11     0.22      1.1     1.31     1.21     0.38     0.49
#> [3,]     0.00     0.00      1.1     1.31     1.21     0.38     0.49
#> [4,]     0.00     0.00      1.1     1.31     1.21     0.38     0.49
#> [5,]     0.00     0.00      1.1     1.31     1.21     0.38     0.49
#>      x6 ~~ x6 x7 ~~ x7 x8 ~~ x8 x9 ~~ x9
#> [1,]     0.36     1.15     0.98     0.92
#> [2,]     0.36     1.15     0.98     0.92
#> [3,]     0.36     1.15     0.98     0.92
#> [4,]     0.36     1.15     0.98     0.92
#> [5,]     0.36     1.15     0.98     0.92
\end{Soutput}
\end{Schunk}

while \textbf{out.reg[[2]]} contains information pertaining to the fit
of each model:

\begin{Schunk}
\begin{Sinput}
round(out.reg[[2]],2)
\end{Sinput}
\begin{Soutput}
#>       lambda conv rmsea     BIC
#>  [1,]   0.00    0  0.19 7805.18
#>  [2,]   0.05    0  0.19 7816.88
#>  [3,]   0.10    0  0.43 9460.55
#>  [4,]   0.15    0  0.43 9460.55
#>  [5,]   0.20    0  0.43 9460.55
#>  [6,]   0.25    0  0.43 9460.55
#>  [7,]   0.30    0  0.43 9460.55
#>  [8,]   0.35    0  0.43 9460.55
#>  [9,]   0.40    0  0.43 9460.55
#> [10,]   0.45    0  0.43 9460.55
#> [11,]   0.50    0  0.43 9460.55
#> [12,]   0.55    0  0.43 9460.55
#> [13,]   0.60    0  0.43 9460.55
#> [14,]   0.65    0  0.43 9460.55
#> [15,]   0.70    0  0.43 9460.55
\end{Soutput}
\end{Schunk}

In this, the user can examine the penalty (lambda), whether the model
converged (``conv''=0), and the fit of each model. By default, two fit
indices are output, both the root mean square error of approximation
(RMSEA; Steiger and Lind 1980), and the Bayesian information criteria
(BIC; Schwarz 1978). Both the RMSEA and BIC take into account the
degrees of freedom of the model, an important point for model selection
in the presence of lasso penalties (and other penalties that set
parameters to zero). Zou, Hastie, and Tibshirani (2007) proved that the
number of nonzero coefficients is an unbiased estimate of the degrees of
freedom for regression. As the penalty increases, select parameters are
set to zero, thus increasing the degrees of freedom, which for fit
indices that include the degrees of freedom in the calculation, means
that although the likelihood of the model may only get worse (increase),
both the RMSEA and BIC can improve (decrease).

Instead of examining the \textbf{out.reg[[1]]} output matrix of
parameter estimates, users also have the option to plot the trajectory
of each of the penalized parameters. This is accomplished with
\textbf{plot_cv(out.reg,pars=1:9)}, resulting in:

\begin{Schunk}
\begin{Sinput}
plot(out.reg,pars=1:9)
\end{Sinput}

\includegraphics{draft1_files/figure-latex/unnamed-chunk-6-1} \end{Schunk}

After a final model (penalty) is chosen, users have the option either
just use the output from \textbf{cv_regsem()}, or the final model can be
re-run with either \textbf{regsem()} or \textbf{multi_optim()} to attain
additional information. In the model above, the best fitting penalty,
according to the BIC, is \(\lambda=0\). However, for demonstration
purposes, we can choose a penalty of 0.2. With this, the model is re-run
with \textbf{regsem()}

\begin{Schunk}
\begin{Sinput}
mod.out <- regsem(out, type="lasso", pars_pen=c(1:9),lambda=0.2)
#summary(mod.out)
\end{Sinput}
\end{Schunk}

Note that the arguments used above correspond to the same arguments in
\textbf{multi_optim()}. However, \textbf{multi_optim()} has additional
optional arguments corresponding to the number or random starts to try.
Additional fit indices can be attained through the
\textbf{fit_indices()} function.

\begin{Schunk}
\begin{Sinput}
fit_indices(mod.out)
\end{Sinput}
\begin{Soutput}
#> $Data_Type
#> [1] "Train"
#> 
#> $fits
#>           Fmin         varFit              p          chisq        p.chisq 
#>        3.35383        0.00000        9.00000     2019.00407        0.00000 
#>           nfac             df           npar              N baseline.chisq 
#>        1.00000       36.00000        9.00000      301.00000      918.85159 
#>    baseline.df           logl            ncp          rmsea    rmsea.lower 
#>       36.00000    -4704.59420        6.61001        0.42850        0.41202 
#>    rmsea.upper     rmsea.pval            CFI            TLI            BIC 
#>             NA        0.00000       -1.24614       -1.24614     9460.55240 
#>            AIC           CAIC         EBIC.5        EBIC.25 
#>     9427.18840     9614.76378     9693.74446     9654.19442
\end{Soutput}
\end{Schunk}

These same fit measures can be accessed through \textbf{cv_regsem}
through changing the defaults with the \textbf{fit.ret=c("rmsea","BIC")}
argument. Finally, instead of assessing these fit indices on the same
sample that the models were run on, a holdout dataset could be used.
This can be done two ways: either with
\textbf{cv_regsem(...,fit.ret2="test")} or with
\textbf{fit_indices(model,CV=TRUE,CovMat=)} and specifying the name of
the holdout covariance matrix.

Structural equation modeling is hard, and pairing with regularization
doesn't make it any easier. Given this, and the number of options
available in the \textbf{regsem} package, a Google group forum was
created in order to answer questions and trouble shoot at
\url{https://groups.google.com/forum/\#!forum/regsem}.

\section{Comparison}\label{comparison}

To compare the different types of penalties in \textbf{regsem}, we
return to the the initial example of the latent growth curve model
displayed in Figure 1. Using the same simulated data, the model can be
run in \textbf{lavaan} as

\begin{Schunk}
\begin{Sinput}
mod1 <- "
i =~ 1*x1 + 1*x2 + 1*x3 + 1*x4
s =~ 0*x1 + 1*x2 + 2*x3 + 3*x4
i ~ c1 + c2 + c3 + c4 + c5 + c6 + c7 + c8 + c9 + c10
s ~ c1 + c2 + c3 + c4 + c5 + c6 + c7 + c8 + c9 + c10
"
lav.growth <- growth(mod1,dat,fixed.x=T)
\end{Sinput}
\end{Schunk}

Comparing different types of penalties in \textbf{regsem} requires a
different specification of the \textbf{type} argument. The options
currently include maximum likelihood (\textbf{"none"}), ridge
(\textbf{"ridge"}), lasso (\textbf{"lasso"}), adaptive lasso
(\textbf{"alasso"}), elastic net (\textbf{"enet"}), SCAD
(\textbf{"scad"}), and MCP (\textbf{"mcp"}). For the elastic net, there
is an additional hyperparameter, \(\alpha\) that controls the tradeoff
between ridge and lasso penalties. This is specified as \textbf{alpha=}
, which has a default of 0.5. Additionally, both the SCAD and MCP have
the additional hyper parameter of \(\gamma\), which is specified as
\textbf{gamma=} and defaults to 3.7 per Fan and Li (2001).

For the purposes of comparison, each of the 20 covariate regressions
were penalized using the lasso, adaptive lasso, SCAD, and MCP, and
compared to the maximum likelihood estimates. In this model, the data
were simulated to have two large effects (both \textit{c1} parameters),
two small effects (both \textit{c2} parameters) and sixteen true zero
effects (\textit{c3-c10} parameters). Note that the covariates were
simulated to have zero covariance among each variable. If there was
substantial collinearity among covariates, the elastic net would be more
appropriate to simultaneously select predictors while also accounting
for the collinearity. The parameter estimates corresponding the the best
fit of the BIC are has the fit of each model, resulting in Table 1,
created using the \textbf{xtable} package (Dahl 2009).

\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
  \hline
 & MLE & lasso & alasso & SCAD & MCP \\ 
  \hline
c1 -$>$ i & 0.92* & 0.72 & 0.91 & 0.94 & 0.92 \\ 
  c2 -$>$ i & 0.07 & 0.00 & 0.00 & 0.00 & 0.00 \\ 
  c3 -$>$ i & 0.10 & 0.00 & 0.00 & 0.00 & 0.00 \\ 
  c4 -$>$ i & 0.07 & 0.00 & 0.00 & 0.00 & 0.00 \\ 
  c5 -$>$ i & 0.04 & 0.00 & 0.00 & 0.00 & 0.00 \\ 
  c6 -$>$ i & -0.25 & 0.00 & 0.00 & 0.00 & -0.19 \\ 
  c7 -$>$ i & 0.11 & 0.00 & 0.00 & 0.00 & 0.00 \\ 
  c8 -$>$ i & -0.13 & 0.00 & 0.00 & 0.00 & 0.00 \\ 
  c9 -$>$ i & -0.03 & 0.00 & 0.00 & 0.00 & 0.00 \\ 
  c10 -$>$ i & 0.09 & 0.00 & 0.00 & 0.00 & 0.00 \\ 
  c1 -$>$ s & 1.18* & 1.09 & 1.22 & 1.24 & 1.24 \\ 
  c2 -$>$ s & 0.29* & 0.19 & 0.28 & 0.35 & 0.35 \\ 
  c3 -$>$ s & 0.18 & 0.09 & 0.00 & 0.00 & 0.00 \\ 
  c4 -$>$ s & -0.08 & 0.00 & 0.00 & 0.00 & 0.00 \\ 
  c5 -$>$ s & -0.18 & 0.00 & 0.00 & 0.00 & 0.00 \\ 
  c6 -$>$ s & 0.25* & 0.00 & 0.00 & 0.00 & 0.00 \\ 
  c7 -$>$ s & -0.18 & -0.04 & 0.00 & 0.00 & 0.00 \\ 
  c8 -$>$ s & 0.26* & 0.10 & 0.00 & 0.00 & 0.00 \\ 
  c9 -$>$ s & -0.06 & 0.00 & 0.00 & 0.00 & 0.00 \\ 
  c10 -$>$ s & 0.08 & 0.00 & 0.00 & 0.00 & 0.00 \\ 
  BIC & 3465.28 & 3427.46 & 3415.05 & 3414.38 & 3417.20 \\ 
   \hline
\end{tabular}
\caption{Parameter estimates for the final models across five estimation methods. Note that * represent significant parameters at p < .05 for maximum likelihood estimation.}
\end{table}

While every regularization method erroneously set both simulated true
intercept effects as zero (non-significant in MLE), both the adaptive
lasso and SCAD correctly identified every true zero effect. The lasso
identified two false effects while the MCP mistakenly identified one.
Additionally, the lasso estimation of the true effects was attentuated
in comparison to the other regularization methods. This is in line with
previous research (Fan and Li 2001), necessitating the use of a two-step
relaxed lasso method (Meinshausen 2007; see Jacobucci, Grimm, and
McArdle 2016) As expected given the small ratio between number of
estimated parameters and sample size, MLE mistakenly identified 3 false
effects as significant.

To compare the performance of each penalization method further,
particularly in the presence of a small parameter to sample size ratio,
a small simulation study was conducted. The same model and effects was
kept, but the sample size was varied to include 80, 200, and 1000 to
demonstrate how MLE improves as sample size increases, while each of the
regularization methods performs well regardless of sample size. Each run
was replicated 200 times. For each regularization method, the BIC was
used to choose a final model among the 40 penalty vales. The results are
displayed in Table 2.

\begin{table}[ht]
\centering
\begin{tabular}{rrrrrrr}
  \hline
 & N & ML & lasso & alasso & SCAD & MCP \\ 
  \hline
 & 80.00 & 0.08 & 0.08 & \textbf{0.04} & 0.05 & 0.20 \\ 
  False Positives & 200.00 & 0.06 & 0.05 & \textbf{0.02} & 0.03 & 0.06 \\ 
   & 1000.00 & 0.05 & 0.08 & \textbf{0.01} & \textbf{0.01} & 0.02 \\ 
   & 80.00 & 0.33 & \textbf{0.31} & 0.35 & 0.35 & 0.31 \\ 
  False Negatives & 200.00 & \textbf{0.19} & \textbf{0.19} & 0.23 & 0.26 & 0.27 \\ 
   & 1000.00 & \textbf{0.00} & \textbf{0.00} & \textbf{0.00} & 0.01 & 0.03 \\ 
   \hline
\end{tabular}
    \caption{Results from the simulation using the model in Figure 1. Each condition was replicated 200 times. False positives represent concluding that the simulated regressions of zero were concluded as nonzero. False negatives are concluding that either the simulated regression values of 1 or 0.2 are in fact zero. Bolded values represent the smallest error per condition.}
\end{table}

For false positives, the adaptive lasso demonstrated the best
performance, where the performance of MLEleveled off at the 0.05 level
at a sample size of 1000 as expected. For false negatives, lasso
penalties demonstrated similar results to MLE. This was expected given
the tendency of the lasso to under-penalize small coefficients in
comparison to the other regularization methods. The adaptive lasso and
SCAD demonstrated slightly worse results, however, outside of the MCP,
each method made either zero or near zero errors at a sample size of
1000. The poor performance of the MCP may be in part due to fixing the
\(\gamma\) penalty to 3.7. Varying this parameter may improve the
performance of the method. In summary, the regularization methods
demonstrated an improvement over maximum likelihood, particularly at
small samples, for a model that had a large number of estimated
parameters.

\section{Discussion}\label{discussion}

This paper provides an introduction to the \textbf{regsem} package,
outlining the mathematical details of regularized structural equation
modeling (RegSEM; Jacobucci, Grimm, and McArdle 2016) and the usage of
the \textbf{regsem} package. RegSEM allows the use of regularization
while keeping the structural equation model intact, adding penalization
directly into the estimation of the model. The application of RegSEM was
detailed using two example models: a latent growth curve model with 20
predictors of both the latent intercept and slope, along with a factor
analysis with one latent factor. With the latent growth curve model, the
small parameter to sample size ratio resulted in a larger number of
false positives in using maximum likelihood estimation. In both the
simulated example and the small simulation, the different types of
regularization in \textbf{regsem} demonstrated better false positive and
negative rates in comparison to maxiumum likelihood across sample sizes.

Broadly speaking, there is a growing amount of research into the
integration between data mining methods and latent variable models.
Specifically, beyond RegSEM, this has taken the form of item response
theory and regularization (Sun et al. 2016), pairing both structural
equation models with decision trees (Brandmaier et al. 2013),
exploratory psychological network analaysis (e.g. Epskamp, Rhemtulla,
and Borsboom 2016), along with many others. The amount of pairing
between methods that have generally been housed in separate camps will
only increase into the future. This type of research will be facilitated
by the general upsurge in the creation of open source software that
gives users a general framework to test models. This was the motivation
behind creating the \textbf{regsem} package, in that users can estimate
models ranging from simple factor analysis models, to latent
longitudinal models with few to many time points, and finally to models
with a large number of latent and observed variables. The use of
regularization allows for the estimation of much larger structural
equation models than before. Specifically, given that latent variables,
as discussed in this article, are more commonly used in the social and
behavioral sciences, sample sizes are generally not large. To estimate
large models with small sample sizes invites increasing amounts of bias
as demonstrated with the simulated data in this paper. Regularization
can be used to reduce the complexity of the model, thus decreasing both
the bias and variance.

With highly constrained structural equation models, achieving model
convergence can be particularly problematic in using \textbf{regsem}.
For instance, with the latent change score model (McArdle and Hamagami
2001), Bayesian regularization methods have less difficulty in reaching
convergence across chains (Jacobucci and Grimm 2017). With the recent
advent of additional sparsity inducing priors, along with new forms of
software such as Stan (Carpenter et al. 2016), for some models it may be
more appropriate to use these Bayesian regularization methods over their
frequentist counterparts. In the realm of Bayesian regularization for
structural equation models, although some research exists (Feng, Wu, and
Song 2017), much more is warranted.

Future research with \textbf{regsem} should focus on a number of
avenues. One is comparing the different forms of regularization,
delineating which method may be best in which setting. Additionally, as
structural equation models become larger, with the advent of much larger
datasets, computational speed will become a principal concern. Although
40 penalties in the models tested above can be run in a matter of
seconds on a standard laptop, larger models can take much longer. To
handle this, future implementation with \textbf{regsem} will test the
inclusion of different types of optimization, specifically testing
whether computation of the Hessian (or approximate Hessian) can reduce
the number of optimization steps.

\subsection{Conclusion}\label{conclusion}

This paper provided a brief overview on the use of the \textbf{regsem}
package as an implementation of regularized structural equation
modeling. Because structural equation modeling encompasses a wide array
of latent variable models, the \textbf{regsem} package was created as a
general package for including different forms of regularization into a
host of latent variable models. RegSEM, and thus the \textbf{regsem}
package, has been evaluated in a wide array of SEM models, including
confirmatory factor analysis (Jacobucci, Grimm, and McArdle 2016),
latent change score models (Jacobucci and Grimm 2017), mediation models
(Serang et al. in press). Future updates to \textbf{regsem} will focus
on decreasing the computational time of large latent variable models in
order to provide an avenue of testing for researchers collecting larger
and larger datasets. RegSEM is a method that operates at all ends of the
data size spectrum: allowing for a reduction in complexity when the
sample size is small, along with dimension reduction in the presence of
large data (both \(N\) and \(P\)).

\bibliography{RJreferences}

\hypertarget{refs}{}
\hypertarget{ref-brandmaier2013}{}
Brandmaier, Andreas M., Timo von Oertzen, John J. McArdle, and Ulman
Lindenberger. 2013. ``Structural Equation Model Trees.''
\emph{Psychological Methods} 18 (1): 71--86.

\hypertarget{ref-browne2001}{}
Browne, Michael W. 2001. ``An Overview of Analytic Rotation in
Exploratory Factor Analysis.'' \emph{Multivariate Behavioral Research}
36 (1). Taylor \& Francis: 111--50.

\hypertarget{ref-carpenter2016stan}{}
Carpenter, Bob, Andrew Gelman, Matt Hoffman, Daniel Lee, Ben Goodrich,
Michael Betancourt, Michael A Brubaker, Jiqiang Guo, Peter Li, and Allen
Riddell. 2016. ``Stan: A Probabilistic Programming Language.''
\emph{Journal of Statistical Software} 20.

\hypertarget{ref-dahl2009xtable}{}
Dahl, David B. 2009. ``Xtable: Export Tables to Latex or Html.'' \emph{R
Package Version}, 1--5.

\hypertarget{ref-donoho1995noising}{}
Donoho, David L. 1995. ``De-Noising by Soft-Thresholding.'' \emph{IEEE
Transactions on Information Theory} 41 (3). IEEE: 613--27.

\hypertarget{ref-epskamp2016generalized}{}
Epskamp, Sacha, Mijke Rhemtulla, and Denny Borsboom. 2016. ``Generalized
Network Psychometrics: Combining Network and Latent Variable Models.''
\emph{ArXiv Preprint ArXiv:1605.09288}.

\hypertarget{ref-fan2001variable}{}
Fan, Jianqing, and Runze Li. 2001. ``Variable Selection via Nonconcave
Penalized Likelihood and Its Oracle Properties.'' \emph{Journal of the
American Statistical Association} 96 (456). Taylor \& Francis: 1348--60.

\hypertarget{ref-feng2017bayesian}{}
Feng, Xiang-Nan, Hao-Tian Wu, and Xin-Yuan Song. 2017. ``Bayesian
Regularized Multivariate Generalized Latent Variable Models.''
\emph{Structural Equation Modeling: A Multidisciplinary Journal}. Taylor
\& Francis, 1--18.

\hypertarget{ref-hastie2015statistical}{}
Hastie, Trevor, Robert Tibshirani, and Martin Wainwright. 2015.
\emph{Statistical Learning with Sparsity: The Lasso and
Generalizations}. CRC Press.

\hypertarget{ref-hoerl1970}{}
Hoerl, Arthur E, and Robert W Kennard. 1970. ``Ridge Regression: Biased
Estimation for Nonorthogonal Problems.'' \emph{Technometrics} 12 (1).
Taylor \& Francis Group: 55--67.

\hypertarget{ref-holzinger1939study}{}
Holzinger, Karl John, and Frances Swineford. 1939. ``A Study in Factor
Analysis: The Stability of a Bi-Factor Solution.'' \emph{Supplementary
Educational Monographs}.

\hypertarget{ref-jacobucci2016package}{}
Jacobucci, Ross. 2017. ``Package 'Regsem'.''
\emph{Https://Cran.r-Project.org/Web/Packages/Regsem/Index.html}.

\hypertarget{ref-jacobucci2016regularized}{}
Jacobucci, Ross, Kevin J Grimm, and John J McArdle. 2016. ``Regularized
Structural Equation Modeling.'' \emph{Structural Equation Modeling: A
Multidisciplinary Journal} 23 (4). Taylor \& Francis: 555--66.

\hypertarget{ref-kline2015principles}{}
Kline, Rex B. 2015. \emph{Principles and Practice of Structural Equation
Modeling}. Guilford publications.

\hypertarget{ref-lam2009sparsistency}{}
Lam, Clifford, and Jianqing Fan. 2009. ``Sparsistency and Rates of
Convergence in Large Covariance Matrix Estimation.'' \emph{Annals of
Statistics} 37 (6B). NIH Public Access: 4254.

\hypertarget{ref-marsh1996assessing}{}
Marsh, Herbert W, and Kit-Tai Hau. 1996. ``Assessing Goodness of Fit: Is
Parsimony Always Desirable?'' \emph{The Journal of Experimental
Education} 64 (4). Taylor \& Francis: 364--90.

\hypertarget{ref-mcardle2005}{}
McArdle, John J. 2005. ``The Development of the Ram Rules for Latent
Variable Structural Equation Modeling.'' \emph{Contemporary
Psychometrics: A Festschrift for Roderick P. McDonald}. Erlbaum. Mahwah,
NJ, 225--73.

\hypertarget{ref-mcardle2001latent}{}
McArdle, John J, and Fumiaki Hamagami. 2001. ``Latent Difference Score
Structural Models for Linear Dynamic Analyses with Incomplete
Longitudinal Data.'' American Psychological Association.

\hypertarget{ref-McArdle_1984}{}
McArdle, John J, and Roderick P. McDonald. 1984. ``Some Algebraic
Properties of the Reticular Action Model for Moment Structures.''
\emph{British Journal of Mathematical and Statistical Psychology} 37
(2). Wiley-Blackwell: 234--51.

\hypertarget{ref-meinshausen2007relaxed}{}
Meinshausen, Nicolai. 2007. ``Relaxed Lasso.'' \emph{Computational
Statistics \& Data Analysis} 52 (1). Elsevier: 374--93.

\hypertarget{ref-meredith1990latent}{}
Meredith, William, and John Tisak. 1990. ``Latent Curve Analysis.''
\emph{Psychometrika} 55 (1). Springer: 107--22.

\hypertarget{ref-statspackage}{}
R Core Team. 2017. \emph{R: A Language and Environment for Statistical
Computing}. Vienna, Austria: R Foundation for Statistical Computing.

\hypertarget{ref-raykov1999desirability}{}
Raykov, Tenko, and George A Marcoulides. 1999. ``On Desirability of
Parsimony in Structural Equation Model Selection.'' \emph{Structural
Equation Modeling: A Multidisciplinary Journal} 6 (3). Taylor \&
Francis: 292--300.

\hypertarget{ref-rosseel2012}{}
Rosseel, Yves. 2012. ``Lavaan: An R Package for Structural Equation
Modeling.'' \emph{Journal of Statistical Software} 48 (2): 1--36.

\hypertarget{ref-schwarz1978estimating}{}
Schwarz, Gideon. 1978. ``Estimating the Dimension of a Model.''
\emph{The Annals of Statistics} 6 (2). Institute of Mathematical
Statistics: 461--64.

\hypertarget{ref-serang2017xmed}{}
Serang, Sarfaraz, Ross Jacobucci, Kim Brimhall, and Kevin J Grimm. in
press. ``Exploratory Mediation Analysis via Regularization.''
\emph{Structural Equation Modeling: A Multidisciplinary Journal}. Taylor
\& Francis.

\hypertarget{ref-steiger1980}{}
Steiger, James H, and John C Lind. 1980. ``Statistically Based Tests for
the Number of Common Factors.'' In \emph{Annual Meeting of the
Psychometric Society, Iowa City, Ia}. Vol. 758.

\hypertarget{ref-sun2016latent}{}
Sun, Jianan, Yunxiao Chen, Jingchen Liu, Zhiliang Ying, and Tao Xin.
2016. ``Latent Variable Selection for Multidimensional Item Response
Theory Models via L\_ \(\{\)1\(\}\) Regularization.''
\emph{Psychometrika} 81 (4). Springer: 921--39.

\hypertarget{ref-thurstone1937}{}
Thurstone, L. L. 1935. \emph{The Vectors of Mind}. Chicago, IL:
University of Chicago Press.

\hypertarget{ref-Tibshirani1996}{}
Tibshirani, Robert. 1996. ``Regression Shrinkage and Selection via the
Lasso.'' \emph{Journal of the Royal Statistical Society. Series B
(Methodological)} 58 (1). Wiley for the Royal Statistical Society:
267--88.

\hypertarget{ref-zhang2010nearly}{}
Zhang, Cun-Hui. 2010. ``Nearly Unbiased Variable Selection Under Minimax
Concave Penalty.'' \emph{The Annals of Statistics} 38 (2). Institute of
Mathematical Statistics: 894--942.

\hypertarget{ref-zou2006adaptive}{}
Zou, Hui. 2006. ``The Adaptive Lasso and Its Oracle Properties.''
\emph{Journal of the American Statistical Association} 101 (476). Taylor
\& Francis: 1418--29.

\hypertarget{ref-zou2005regularization}{}
Zou, Hui, and Trevor Hastie. 2005. ``Regularization and Variable
Selection via the Elastic Net.'' \emph{Journal of the Royal Statistical
Society: Series B (Statistical Methodology)} 67 (2). Wiley Online
Library: 301--20.

\hypertarget{ref-zou2006sparse}{}
Zou, Hui, Trevor Hastie, and Robert Tibshirani. 2006. ``Sparse Principal
Component Analysis.'' \emph{Journal of Computational and Graphical
Statistics} 15 (2). Taylor \& Francis: 265--86.

\hypertarget{ref-Zou2007}{}
---------. 2007. ``On the Degrees of Freedom of the Lasso.'' \emph{The
Annals of Statistics} 35 (5). Institute of Mathematical Statistics:
2173--92.
doi:\href{https://doi.org/10.1214/009053607000000127}{10.1214/009053607000000127}.

\address{%
Ross Jacobucci\\
University of Notre Dame\\
118 Haggar Hall, Notre Dame, IN 46556\\
}
\href{mailto:rcjacobuc@gmail.com}{\nolinkurl{rcjacobuc@gmail.com}}